---
title: "Inference, Etc"
author: "Preye Indiamaowei"
format: html
embed-resources: true
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
```

## Problem 1
The first two problems in this set refer to the ISLR2::Carseats set, which includes simulated sample data for 400 stores selling a particular brand of child carseat. We view this data as a simple random sample from a larger population.

Does this data provide evidence against the company’s claim that the average price of their own seats is $110? Follow all best practices from class, including hypotheses, a relevant visualization, and a conclusion in non-technical language.

```{r}
library(ISLR2)
glimpse(Carseats)
```

```{r}
avg_price = mean(Carseats$Price, na.rm = TRUE)
```
The mean price according to this data is $115.80.

```{r}
Carseats %>% 
  ggplot(aes(x = Price)) +
  geom_histogram() +
  geom_vline(xintercept = 110)
```
Null hypothesis: 110 ($H_0 : \mu = 110 $)

Alternative hypothesis: mu isn't 110
$(H_a : \mu \ne 110)$

```{r}
t.test(Carseats$Price,
mu = 110)
```
the p-value of 0.000001429 means that data like ours will only occur .00001429% of the time when we gather samples of this size. The null hypothesis does not seem to be compatible with the data. We have good evidence against $H_0$, so we reject it and can conclude $\mu \ne 110$.

We can conclude the average price of this company's car seats is not $110.


## Problem 2
Compute a level 90% confidence interval for the mean competitor price in the population from which this sample is drawn. What is the sample mean? What is the margin of error?
```{r}

```


## Problem 3
Watch my video on contingency tables. Use these tools to create a table with counts of penguin species and home islands in the penguins set. Include marginal counts in your table.
```{r}
glimpse(penguins)
```
```{r}
# species, island
penguins_table <- table(penguins$species, penguins$island)

addmargins(penguins_table)

penguins_table
```


## Problem 4
Push your solutions to this set to your GitHub page, including both the Quarto source and the rendered html. Use file names of the form, “math250_nov5_yourname.”

Provide a clickable link to the repository on your personal GitHub where these files can be found.