---
title: "Practice with ggplot"
author: "Preye Indiamaowei"
date: "8 September 2025"
format: html
embed-resources: true

---

## Problem 1

We will be working with the `attrition` data set from the `modeldata` package. Let's take a look at the data set. 
```{r}
#| message: false

library(tidyverse)
library(modeldata)
glimpse(attrition)
```
Now let's use the `ggplot()` function to create a box plot from the data set.

```{r}
ggplot(attrition, aes(x = MonthlyIncome,
                      y = Education,
                      fill = Education)) +
  geom_boxplot() +
  labs(x = " Monthly Income",
       y = "Education",
       title = "Learn to Earn") +
  scale_fill_brewer(palette = "Dark2")
```
Note: I couldn't figure out how to get the boxes to be different colors.

## Problem 2

Next, we will create a bar chart to look at the `Age` variable.

```{r}
ggplot(attrition, aes(x = Age)) +
  geom_bar(color = "darkgreen",
           fill = "lightgreen") +
  theme_minimal()
```

## Problem 3

Next, we will generate an accessible bar chart stressing the different levels of job satisfaction, while also showing the relative attrition rates within each group.

```{r}
ggplot(attrition, aes(x = JobSatisfaction,
                      fill = Attrition)) +
  geom_bar() +
  labs(x = "Job Satisfaction",
       y = "Frequency",
       title = "Attrition and Job Satisfaction") +
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal()
```

## Problem 4

Next, we will generate an accessible bar chart clearly comparing the different attrition rates for various levels of job satisfaction.

```{r}
ggplot(attrition, aes(fill = JobSatisfaction,
                      x = Attrition)) +
  geom_bar() +
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal()
```

## Problem 5

Finally, we will create a bar chart that displays the monthly income of different education fields.

```{r}
ggplot(attrition, aes(x = MonthlyIncome,
                      fill = EducationField)) +
  geom_histogram()+
  facet_wrap(~EducationField) +
  labs(title = "Income by Field",
       x = "Monthly Income",
       caption = "Source: IBM Watson Analytics Lab") +
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal()
```

