---
title: "Class Notes Template"
author: "Preye Indiamaowei"
format: html
embed-resources: true
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
```

## Lists

A list is a generalization of the data frame.

```{r}
my_list <- list(nums = 1:5,
                vars = c("area",
                         "peri",
                         "shape",
                         "perm"),
                data = rock)

str(my_list) # see the structure of the object
```

Extract elements using `$` or double brackets.

```{r}
my_list$nums
my_list[["vars"]]
```
## Repetition

Loops exist in R, but we don't use them. Instead, we usually can just apply a specified function to every element of a vector, data frame, or list.

Example of how this can get complicated:
```{r}
typeof(my_list)
```

I'd like to apply `typeof()` to every element of the list. Gard reccomends `map` and its friends.
```{r}
map(my_list, typeof)
```
This always gives back a list. We can specify the sort of object we want back by using a related function.
```{r}
map_chr(my_list, typeof)
```

General rule: If you need to so the same thing 3 times, consider automating it.

What if the thing we need to automate isn't just a function like `typeof()`? For instance, suppose I want to keep the four lowest values of each column of the `rock` set?

```{r}
glimpse(rock)
```

## Writing Functions

A function to keep the four smallest values of a numerical vector:
```{r}
keep_smallest <- function(nums){
  sort(nums)[1:4]
}

keep_smallest(rock$area)

map(rock, keep_smallest)
```

** Functions are first-class objects. We create them using `function()`, a constructor.

