---
title: "R Notes 9/19/25"
author: "Preye Indiamaowei"
format: html
embed-resources: true
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(modeldata)

```

Prof Gard's Plot
```{r}
#| warning: false

ggplot(scat, aes(y = Species,
                 x = d15N,
                 fill = Species)) +
  geom_boxplot(outlier.shape = NA,
               show.legend = FALSE) +
  
  

```
There are two missing values which I haven't explored.

## Arranging a data set by a variable

Let's arrange scat by `d15N`.

```{r}
scat %>%
  select(Species, 
         Site, 
         d15N) %>% 
  arrange(desc(d15N)) %>% 
  head() # show just the first few rows
```

We can arrange by multiple variables, including categorical ones.

```{r}
scat %>%
  select(Species, 
         Site, 
         d15N,
         Location) %>% 
  arrange(Location, d15N) %>% 
  print(n = 110)
```

# Adding/modifying columns

Use the `mutate` command. Let's get the ratio of `diameter` to `length` in the `scat` set.

```{r}
scat %>% 
  select(Species, 
         Diameter,
         Length) %>% 
  mutate(Shape = Diameter / Length) %>% 
  head()
```

We can also use this command to change an existing column.

Last comment: `arrange` always gives back a set with the same dimentions as the original. `mutate` always has the same number of rows but psiibly have more columns. `filter`has the same number of columns, and `select` has the same number of rows.

