---
title: "R notes 10/20"
format: html
embed-resources: true
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(gt)

msleep1 <- data.frame(name = c("Cheetah", 
                               "Dog"),
                      genus = c("Acinonyx",
                                "Canis"))
msleep2 <- data.frame(name = c("Cow",
                               "Goat"),
                      genus = c("Bos",
                                "Capri"))
msleep_vore <- data.frame(name = c("Cheetah", 
                                   "Cow",
                                   "Goat",
                                   "Brown bat"),
                          vore = c("carni",
                                   "herbi",
                                   "herbi",
                                   "insecti"))
```

## Combining sets: new observations, same variables

```{r}
msleep1
msleep2
```

Here the variable names match up. All we need to do is copy/paste.

```{r}
msleep1 %>% 
  bind_rows(msleep2)
```

Check out the help file with `?bind_rows`.

```{r}
msleep_genus <- msleep1 %>% 
  bind_rows(msleep2,
            .id = "Researcher")
msleep_genus
```

## Combining sets: different variables

I'd like to add `vore` informaion to `msleep_genus`. We do this with `left_join`.

```{r}
msleep_genus %>% 
  left_join(msleep_vore,
            by = "name")
```

Order matters!
```{r}
msleep_vore %>% 
  left_join(msleep_genus,
            by = "name")
```
If you leave put `by`, then R will join by all the matching columns. This is called a *natural join*. 

```{r}
msleep_vore %>% 
  left_join(msleep_genus)
```

Prof Gard does not reccomend this. It can go wrong in a million different ways.

## Filtering Joins

We can also use a second data set to filter our set, for instance, only keeping observations that also appear in the second set (`semi_join`) or that don't appear in the second (`anti-join`).