---
title: "R Notes 9/26/25"
author: "Preye Indiamaowei"
format: html
embed-resources: true
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(dplyr)
```

## Even more on grouping

We can group by variables that we define on the fly.

```{r}
diamonds %>% 
  group_by(big = carat > median(carat)) %>% 
  summarize(avg_price = mean(price))
```

We can group by more than one variable at a time. 

```{r}
diamonds %>% 
  group_by(big = carat > median(carat),
           cut) %>% 
  summarize(avg_price = mean(price))
```

We can strip away the grouping with `ungroup` if we need.

Order matters:
```{r}
diamonds %>% 
  group_by(cut,
           big = carat > median(carat)) %>% 
  summarize(avg_price = mean(price))
```

## Slicing a set

What are the five largest diamonds? What are the five largest of each cut?

```{r}
diamonds %>% 
  slice_max(order_by = carat,
            n = 5)
```

Notice we still have all the variables.

```{r}
diamonds %>% 
  group_by(cut) %>% 
    slice_max(order_by = carat,
              n = 5)
```

## Grouped mutates & filters

Let's only keep diamonds that are larger than the median for their cut.

```{r}
diamonds %>% 
  group_by(cut) %>% 
  filter(carat > median(carat))
```

We can also add columns groupwise using mutate.

```{r}
diamonds %>% 
  group_by(cut) %>% 
  mutate(median_carat_by_cut = carat > median(carat))
```

```{r}
diamonds %>% 
  group_by(cut) %>% 
  mutate(median_carat_by_cut = carat > median(carat)) %>% 
  select(carat, cut, color, clarity, depth, table, price)
```

## Alt text for plots

We add alt text to plots using a chunk option.

```{r}
#| fig-alt: A bar chart showing higher counts for better cuts of diamonds

ggplot(diamonds, aes(x = cut)) +
  geom_bar()
```

