---
title: "R Notes 9/8/25"
author: "Preye Indiamaowei"
date: "8 September 2025"
format: html
embed-resources: true

---

If you have `editor: visual` in your YAML, please delete it.

## Bar charts

Bar charts are used to visualize 1-2 categorical variables. We'll explore this using the `Carseats` data set from the `ISLR2` package.

```{r}
#| message: false


library(ISLR2)
library(tidyverse)

```

Use the `data()` command to view all available datasets.

```{r}
glimpse(Carseats)
```

Let's consider `ShelveLoc` and `US`

```{r}
ggplot(Carseats, aes(x = ShelveLoc,
                     fill = US)) +
  geom_bar() +
  scale_fill_brewer(palette = "Dark2")
```

This plot stresses the relative counts of the `ShelveLoc` variable. It also shows `US`, but the relative proportions aren't so clear. We can flip the variables to reverse this.

```{r}
ggplot(Carseats, aes(fill = ShelveLoc,
                     x = US)) +
  geom_bar() +
  scale_fill_brewer(palette = "Dark2")
```

Let's stress the relative proportions.
Same data, same aesthetics.

```{r}
ggplot(Carseats, aes(fill = ShelveLoc,
                     x = US)) +
  geom_bar(position = "fill") +
  scale_fill_brewer(palette = "Dark2")
```
It seems that stores in the US have slightly more "good" shelving than stores outside the US. The bad shelving count seems to be about the same proportionally.

Check out the help file for more options. `?geom_bar`.

Finally, we can use `position = "dodge"` to get a compromise.

```{r}
ggplot(Carseats, aes(fill = ShelveLoc,
                     x = US)) +
  geom_bar(position = "dodge") +
  scale_fill_brewer(palette = "Dark2")
```
